---
title: "Testing Transformation Methods"
output: 
  html_document: 
    toc: yes
---

```{r, message = FALSE}
library(pmartRmems)
library(ggplot2)
library(tidyverse)
library(ALDEx2)
```


```{r}
data("rRNA_obj")
mice <- rRNA_obj
```


# Mice Data

We'll run the transformations for all the methods and then compare their distributions at the bottom. The dafult "shift" value of 0.5 is used.

### CLR method

```{r}
mice_clr <- transform_data(mice, method = "clr")
```

By default a message is printed showing the indices of all rows of zeros that were removed. Optional arguments include <code>quiet = TRUE</code> to supress this message, or <code>include_zero_rows = TRUE</code> to refrain from removing rows with all zeros. 

### ALR Method

The function returns an error if the ALR method is used but a basis OTU isn't chosen.

```{r, error = TRUE}
mice_alr <- transform_data(mice, method = "alr", quiet = TRUE)
mice_alr <- transform_data(mice, method = "alr", basis_ind = 1, quiet = TRUE)
```

### IQLR Method

```{r}
mice_iqlr <- transform_data(mice, method = "iqlr", quiet = TRUE)
```

### Zero Method

```{r, error = TRUE}
mice_zero <- transform_data(mice, method = "zero", quiet = TRUE)
names(mice$f_data)
mice2 <- group_designation(mice, "Treatment")
mice_zero <- transform_data(mice2, method = "zero", quiet = TRUE)
```

### User Method

As with the ALR method, the user must give one or more indices for basis OTUs.

```{r, error = TRUE}
mice_user <- transform_data(mice, method = "user", quiet = TRUE)
mice_user <- transform_data(mice, method = "user", basis_ind = 1:10, quiet = TRUE)
```


## Method Comparison

Here the raw data is compared with the five transformation methods. Odd patterns are manifest in the ALR and User methods likely due to random OTUs being chosen for a basis rather than choosing them based on properties of the data.

```{r, echo = FALSE}
gather_rrna <- function(rRNA_obj, method) {
  rRNA_obj$e_data %>% gather("Sample", "Reads", 2:ncol(rRNA_obj$e_data)) %>% mutate(Method = method)
}
```

```{r, echo = FALSE}
datlist <- list(raw = mice, clr = mice_clr, iqlr = mice_iqlr, 
                alr = mice_alr, user = mice_user, zero = mice_zero)
plotdat <- list(df = NULL)
for (i in 1:6) {
  newdat <- gather_rrna(datlist[[i]], names(datlist)[i])
  plotdat$df <- rbind(plotdat$df, newdat)
}

plotdat$df$Facet = factor(plotdat$df$Method, levels = unique(plotdat$df$Method))
```

```{r, fig.width = 12, fig.height = 8, echo = FALSE}
ggplot(plotdat$df) + geom_boxplot(aes(Sample, Reads), fill = "blue") +
  facet_wrap(~ Facet, scales = "free_y") + theme_grey(base_size = 18) + theme(axis.text.x=element_blank())
```


## OTUs

```{r, fig.height = 24, fig.width = 12}
# mice_long <- dat %>% gather("Sample", "Reads", 2:ncol(dat)) %>% rename_(OTU = OTU_ID) %>% 
#   mutate(Reads = log(Reads + 1), OTU = factor(OTU, levels = unique(dat[,OTU_ID])))
temp <- subset(plotdat$df, Method != "raw")
rng <- range(temp$Reads)
mn <- floor(rng[1])
mx <- ceiling(rng[2])
plotfun <- function(method) {
  sub <- subset(plotdat$df, Method == method)
  p <- ggplot(sub) + geom_raster(aes(OTU.ID, Sample, fill = Reads)) + 
    theme(axis.text.x  = element_blank(),
          axis.text.y  = element_blank()) +
    ggtitle(method)

  if(method != "raw") {
    p <- p + scale_fill_gradientn(colours = c("blue","green","red"), 
                         values = scales::rescale(c(mn, mn/10, 0, mx/10, mx)),
                         limits=c(mn, mx)) +
                         labs(x = "OTU")
  } else {
    p <- p + scale_fill_gradientn(colours = c("green","blue","red"), 
                         values = scales::rescale(c(0, 1, 100, 1000, 10000, 
                                                    100000, max(plotdat$df$Reads)))) +
                         labs(x = "OTU")
  }
  
  return(p)
}

plotlist <- lapply(names(datlist), plotfun)
cowplot::plot_grid(plotlist[[1]], plotlist[[2]], plotlist[[3]], 
                   plotlist[[4]], plotlist[[5]], plotlist[[6]], ncol = 1)
```





# Soil Data

```{r}
data("rRNA_obj2")
soil <- rRNA_obj2

soil_clr <- transform_data(soil, method = "clr")
soil_iqlr <- transform_data(soil, method = "iqlr")
soil_alr <- transform_data(soil, method = "alr", basis_ind = 1)
soil_user <- transform_data(soil, method = "user", basis_ind = 1:10)
names(soil$f_data)
soil2 <- group_designation(soil, "Treatment")
soil_zero <- transform_data(soil2, method = "zero")
```

```{r, echo = FALSE}
datlist <- list(raw = soil, clr = soil_clr, iqlr = soil_iqlr, 
                alr = soil_alr, user = soil_user, zero = soil_zero)
plotdat <- list(df = NULL)
for (i in 1:6) {
  newdat <- gather_rrna(datlist[[i]], names(datlist)[i])
  plotdat$df <- rbind(plotdat$df, newdat)
}

plotdat$df$Facet = factor(plotdat$df$Method, levels = unique(plotdat$df$Method))
```

```{r, fig.width = 12, fig.height = 8, echo = FALSE}
ggplot(plotdat$df) + geom_boxplot(aes(Sample, Reads), fill = "blue") +
  facet_wrap(~ Facet, scales = "free_y") + theme_grey(base_size = 18) + theme(axis.text.x=element_blank())
```


# Selex Data

```{r}
data("selex")
# convert to rRNAdata object
e_data <- selex %>% mutate(OTU = row.names(selex)) %>% select(OTU, everything())
f_data <- data.frame(SampleID = colnames(selex), 
                     Condition = c(rep("NonSelected", 7), rep("Selected", 7)))
selex_rrna <- as.rRNAdata(e_data, f_data, edata_cname = "OTU", fdata_cname = "SampleID")

selex_clr <- transform_data(selex_rrna, method = "clr")
selex_iqlr <- transform_data(selex_rrna, method = "iqlr")
selex_alr <- transform_data(selex_rrna, method = "alr", basis_ind = 1)
selex_user <- transform_data(selex_rrna, method = "user", basis_ind = 1:10)
names(selex_rrna$f_data)
selex2 <- group_designation(selex_rrna, "Condition")
selex_zero <- transform_data(selex2, method = "zero")
```

```{r, echo = FALSE}
datlist <- list(raw = selex_rrna, clr = selex_clr, iqlr = selex_iqlr, 
                alr = selex_alr, user = selex_user, zero = selex_zero)
plotdat <- list(df = NULL)
for (i in 1:6) {
  newdat <- gather_rrna(datlist[[i]], names(datlist)[i])
  plotdat$df <- rbind(plotdat$df, newdat)
}

plotdat$df$Facet = factor(plotdat$df$Method, levels = unique(plotdat$df$Method))
```

```{r, fig.width = 12, fig.height = 8, echo = FALSE}
ggplot(plotdat$df) + geom_boxplot(aes(Sample, Reads), fill = "blue") +
  facet_wrap(~ Facet, scales = "free_y") + theme_grey(base_size = 18) + theme(axis.text.x=element_blank())
```






